---
title: "ScholarData"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(knitr)
```

```{r}
sd <- read_csv("data/applicant_countries_rfp16-18.csv")

developing_countries <- c("Other", "Bangladesh", "India", "Brazil", "Mexico", "Romania", "Costa Rica", "Chile", "Bolivia", "Ecuador", "Poland", "Indonesia", "Argentina", "South Africa", "Ukraine", 
                          "Bhutan", "Colombia", "Cameroon", "Nepal", "Malawi", "Malaysia", "Pakistan", "Nigeria", "Namibia", "Papa New Guinea", "Thailand", "Virgin Islands, U.s.", "Cote D'Ivoire",
                          "Uganda", "Tanzania, United Republic of", "Ghana", "Singapore", "Guatemala")
firstworld_countries <- c("United States", "Germany", "United Kingdom", "Canada", "Australia", "Austria", "Portugal", "Hungary", "Italy", "Finland", "Denmark", "New Zealand")
western_countries <- c("United States", "Germany", "United Kingdom", "Canada", "Austria", "Portugal", "Hungary", "Italy", "Finland", "Denmark")

sd_aw <- sd %>% 
  filter(Status == "Awarded") %>% 
  #add column that indicates if project was conducted from country of awardee's nationality
  rename(project_country = `Project Country`) %>% 
  mutate(project_nat = ifelse(project_country == Nationality & Nationality == Country & project_country == Country, "all_yes", 
                              ifelse(project_country == Nationality & Nationality != Country & project_country != Country, "project_nat_yes",
                                     ifelse(project_country != Nationality & Nationality == Country & project_country != Country, "nat_country_yes",
                                            ifelse(project_country != Nationality & Nationality != Country & project_country == Country, "project_country_yes",
                                                   ifelse(project_country != Nationality & Nationality != Country & project_country != Country, "all_no", NA))))),
         usa_uk = ifelse(Country %in% c("United States", "United Kingdom") & !project_country %in% c("United States", "United Kingdom"), "working_abroad", "working_home"))

sd_projSum <- sd_aw %>%
  mutate(tot_n = n()) %>% 
  group_by(project_nat, tot_n) %>% 
  summarise(n = n()) %>% 
  mutate(percent = round(n/tot_n, 2) * 100)

sd_usSum <- sd_aw %>%
  mutate(tot_n = n()) %>% 
  group_by(usa_uk, tot_n) %>% 
  summarise(n = n()) %>% 
  mutate(percent = round(n/tot_n, 2) * 100)

all_yes <- sd_projSum %>% filter(project_nat == "all_yes") %>% ungroup() %>% dplyr::select(percent)
project_nat_yes <- sd_projSum %>% filter(project_nat == "project_nat_yes") %>% ungroup() %>% dplyr::select(percent)
working_abroad <- sd_usSum %>% filter(usa_uk == "working_abroad") %>% ungroup() %>% dplyr::select(percent)

#of the 16/35 students working in their country of origin, how many are NOT from US or European countries?
sd_dev <- sd_aw %>%
  filter(project_nat == "all_yes") %>% 
  filter(!Country %in% western_countries)

all_yes_devcountry <- nrow(sd_dev)
```

```{r}
projSum <- sd_projSum %>%
  mutate(percent_formatted = paste(percent, "% ", "(", n, "/", tot_n, ")", sep = "")) %>% 
  dplyr::select(project_nat, percent_formatted) %>% 
  spread(key = project_nat, value = percent_formatted)

kable(projSum, col.names = c("Project!=Nationality!=Country", "Project=Nationality=Country", "Project!=Nationality=Country", "Project=Country!=Nationality", "Project=Nationality!=Country"))
```



Of the 35 students we've funded over the past 3 years, `r all_yes`% are from the country where the project is occurring and is studying in that country. An additional `r project_nat_yes`% students have projects in their country of nationality but are attending university elsewhere (typically the United States). `r working_abroad`% are US or UK students working abroad in developing countries.

Of the 16/35 students working in their country of origin that are also studying in their country of origin, `r all_yes_devcountry` are not from US or European countries.